HyperLogLog: Exponentially Bad in Adversarial Settings

### 2021-08-28 docs/HLL_exp_bad/main.tex
more work on formal analysis to improve presentation

\newif\ifIEEEversion % turn on IEEE style 
\IEEEversionfalse
\newif\ifanon % needed for acks, suppression of authors, etc
\anonfalse

@@

\subsection{Intuition for Formal Security Analysis}

@@

We next describe the simulator ${\mathcal{S}}$ that we will use in our further analysis: ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ in a blackbox manner and handling all of its queries as follows: 
\begin{itemize}
\item \texttt{init($\cdot$)}: on input $n$, $\ell$, $\alpha_m$, ${\mathcal{S}}$ initialises an HLL sketch as in Figure~\ref{fig:hll}; that is, it creates an array of $m=2^n$ buckets $M[j]$ and initialises them all to 0.
\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${\mathcal{S}}$ ignores $x$, sets $y \leftarrow_{\$} \{0,1\}^{n+\ell}$, and proceeds to update the buckets using this value of $y$ as in the original HLL algorithm in Figure~\ref{fig:hll}.
\item \texttt{read}:  ${\mathcal{S}}$ returns the HLL cardinality estimate computed as per Figure~\ref{fig:hll}.
\item \texttt{corrupt}: ${\mathcal{S}}$ returns the current vector $(M[j])_{j=1}^{m}$ of HLL bucket contents that it has stored.
\end{itemize}
At the end of ${\mathcal{A}}$'s execution, ${\mathcal{S}}$ takes ${\mathcal{A}}$'s output and forwards it to ${\mathcal{E}}$ as its own output. Notice that the running time of  ${\mathcal{S}}$ is (roughly) equal to that of ${\mathcal{A}}$.

@@

We require one further standard definition.
\begin{definition}
We define the PRF advantage of an adversary ${{\mathcal{B}}}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$). Thus: 
\[
\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) := \lvert 2\Pr[d' = d] - 1 \rvert\,.
\]
\end{definition}
Here, we assume without loss of generality that all of ${{\mathcal{B}}}$'s queries are distinct. See~\cite[Section 4.4.1]{BonehShoup} for details of the PRF security definition.

@@

>>>>

\newif\ifIEEEversion % turn on IEEE style 
\IEEEversiontrue
\newif\ifanon % needed for acks, suppression of authors, etc
\anontrue

@@

\subsection{Intuition for Formal Security Analysis}\label{sec:informal}

@@

\vspace{0.1cm}
\noindent\textbf{A specific simulator ${\mathcal{S}}$.}
We next describe the simulator ${\mathcal{S}}$ that we will use in our further analysis: ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ in a blackbox manner and handling all of its queries as follows: 
\begin{itemize}
\item \texttt{init($\cdot$)}: on input $n$, $\ell$, $\alpha_m$, ${\mathcal{S}}$ initialises an HLL sketch as in Figure~\ref{fig:hll}; that is, it creates an array of $m=2^n$ buckets $M[j]$ and initialises them all to 0.
\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${\mathcal{S}}$ ignores $x$, sets $y \leftarrow_{\$} \{0,1\}^{n+\ell}$, and proceeds to update the buckets using this value of $y$ as in the original HLL algorithm in Figure~\ref{fig:hll}.
\item \texttt{read}:  ${\mathcal{S}}$ returns the HLL cardinality estimate computed as per Figure~\ref{fig:hll}.
\item \texttt{corrupt}: ${\mathcal{S}}$ returns the current vector $(M[j])_{j=1}^{m}$ of HLL bucket contents that it has stored.
\end{itemize}
At the end of ${\mathcal{A}}$'s execution, ${\mathcal{S}}$ takes ${\mathcal{A}}$'s output and forwards it to ${\mathcal{E}}$ as its own output. Notice that the running time of  ${\mathcal{S}}$ is (roughly) equal to that of ${\mathcal{A}}$.

@@

We require one further standard definition.
\begin{definition}
Let $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
We define the PRF advantage of an adversary ${{\mathcal{B}}}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$). We write: 
\[
\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) := \lvert 2\Pr[d' = d] - 1 \rvert\,.
\]
\end{definition}
Here, we assume without loss of generality that all of ${{\mathcal{B}}}$'s queries are distinct. See~\cite[Section 4.4.1]{BonehShoup} for details of the PRF security definition.

@@

This result tightly relates the security of HLL-$F$ against cardinality manipulation to that of the PRF security of the function $F$ used in its construction (for the reasons explained in Section~\ref{sec:informal}).
